PineAPPL:\\ combining EW and QCD corrections for PDF fits

### 2020-08-08 paper/paper.tex
Sect 3 replaced
Title
adding preprint number
Added Nikhef preprint number
small tweaks+comments
Added tabularx package
Fix more Latex warnings

\documentclass[a4paper,11pt]{article}
\pdfoutput=1

\usepackage[UKenglish]{babel}
\usepackage{booktabs}
\usepackage{csquotes}
\usepackage{jheppub}
\usepackage[T1]{fontenc}
\usepackage{listings}
\usepackage{lmodern}
\usepackage[binary-units=true,detect-weight=true,table-format=4.2]{siunitx}
\usepackage{xcolor}
\usepackage[percent]{overpic}
\usepackage{mathcomp}

\newcommand{\alphas}{\alpha_\mathrm{s}}

\title{PineAPPL:\\ EW and mixed QCD--EW corrections for PDF fits}

@@

\abstract{We introduce \textsc{PineAPPL}, a computer library that makes it possible to produce fast-interpolation 
grids for fitting parton distribution functions (PDFs)
including corrections of strong and electroweak origin. After presenting
the features and performance of such a code, we show how to interface \textsc{PineAPPL} with general-purpose
softwares that allow the user to compute such corrections, providing an explicit example with
\textsc{MadGraph5\_aMC@NLO}. We showcase some phenomenological result focusing on processes and observables
relevant for PDF fits, and 
we discuss what steps need to be undertaken in the presentation of experimental data in order to have them
consistently employed in a fit including electroweak corrections.}

@@

\input{sec-introduction}
\input{sec-pineappl}
\input{sec-results-marco}
\input{sec-doublecounting}
\input{sec-conclusion}

>>>>

\documentclass[a4paper,11pt]{article}
\pdfoutput=1
\pdfsuppresswarningpagegroup=1

\usepackage[UKenglish]{babel}
\usepackage{booktabs}
\usepackage{csquotes}
\usepackage{jheppub}
\usepackage[T1]{fontenc}
\usepackage{listings}
\usepackage{lmodern}
\usepackage[binary-units=true,detect-weight=true,table-format=4.2]{siunitx}
\usepackage{xcolor}
\usepackage[percent]{overpic}
\usepackage{mathcomp}
\usepackage{tabularx}

\newcommand{\alphas}{\alpha_\mathrm{s}}
\preprint{Nikhef-2020-026 TIF-UNIMI-2020-22}

\title{PineAPPL:\\ combining EW and QCD corrections for PDF fits}

@@

\abstract{We introduce \textsc{PineAPPL}, a computer library allowing to produce fast-in\-ter\-po\-la\-tion
grids for fitting parton distribution functions (PDFs)
including corrections of strong and electroweak origin. After presenting
the features and performance of such a code, we show how to interface \textsc{PineAPPL} with general-purpose
softwares that allow the user to compute such corrections, providing an explicit example with
\textsc{MadGraph5\_aMC@NLO}. We showcase some phenomenological result focusing on processes and observables
relevant for PDF fits, and 
we discuss what steps need to be undertaken in the presentation of experimental data in order to have them
consistently employed in a fit including electroweak corrections.}

@@

\input{sec-introduction}
\input{sec-pineappl}
\input{sec-results}
\input{sec-doublecounting}
\input{sec-conclusion}

### 2020-08-08 paper/sec-results.tex
small tweaks+comments
Fix more Latex warnings

In this section we demonstrate the capabilities of \textsc{PineAPPL} by
computing fast-interpolation grids, accurate to NLO QCD and NLO QCD+EW,
for a representative set of processes in which EW corrections may have a
sizeable effect on the accuracy of the theoretical predictions.
In order to consider realistic kinematics for these
processes, we rely on the measurements that are commonly included in PDF
fits. Our aim is twofold. First, we want to validate the results
obtained with \textsc{PineAPPL}; second, we want to assess the
impact of the EW corrections for the specific experimental setups. We start by describing the
setup employed for the computation, then we discuss the features of each process, along with
the corresponding results.

@@

We generate each process by means of the Universal FeynRules Output
(UFO)~\cite{Degrande:2011ua} model {\tt loop\_qcd\_qed\_sm\_Gmu},
included as standard in {\sc MG5\_aMC}. It contains the UV and $R_2$
counterterms relevant to NLO QCD and EW corrections, the latter in the
$\overline{G}_\mu$ scheme. The model features five massless quark flavours,
sets the CKM matrix equal to the identity, and is compatible with the usage of
the complex mass (CM) scheme~\cite{Denner:1999gp,Denner:2005fg} for all massive particles, see
ref.~\cite{Frederix:2018nkq} for details. We use this scheme
for all processes that involve only massless particles in the final state.
The photon is always considered as part of the proton in the initial state and
of any hadronic jet produced in the final state: 
 in particular, PI contributions are consistently included at LO and NLO.\footnote{We employ the $\overline{G}_\mu$ scheme also for the QED coupling entering vertices involving initial-state photons, see section~4.3.3 of ref.\cite{Denner:2019vbn}.}
We use a PDF set that contains a photon PDF, namely
{\tt NNPDF31\_nlo\_as\_0118\_luxqed}~\cite{Bertone:2017bme}. We evaluate the PDF
uncertainty associated to the theoretical predictions a posteriori, that is,
we convolve the fast-interpolation grid generated with {\sc PineAPPL} with
each member in the PDF set, and we compute the associated standard deviation.

@@

%-------------------------------------------------------------------------------
\begin{figure}[!p]
    \centering
    \includegraphics[width=0.5\textwidth]{figures/pineappl_CMSDY2D11_bin3}%
    \includegraphics[width=0.5\textwidth]{figures/pineappl_CMSDY2D11_bin4}\\
    \includegraphics[width=0.5\textwidth]{figures/pineappl_CMSDY2D11_bin5}%
    \includegraphics[width=0.5\textwidth]{figures/pineappl_CMSDY2D11_bin6}\\
    \caption{Same as figure~\ref{fig:atlaszhighmass49fb}, but for the CMS
      double-differential Drell--Yan lepton pair measurement at a
      centre-of-mass energy of \SI{7}{\tera\electronvolt}~\cite{Chatrchyan:2013tia}. Displayed are
      only two of the six invariant mass bins available, respectively below the
      Z-boson mass peak, $\SI{45}{\giga\electronvolt}<M_{\ell\bar\ell}<\SI{60}{\giga\electronvolt}$, on the Z-boson mass
      peak, $\SI{60}{\giga\electronvolt}<M_{\ell\bar\ell}<\SI{120}{\giga\electronvolt}$, above the mass peak,
      $\SI{120}{\giga\electronvolt}<M_{\ell\bar\ell}<\SI{200}{\giga\electronvolt}$, and at very high invariant masses,
      $\SI{200}{\giga\electronvolt}<M_{\ell\bar\ell}<\SI{1500}{\giga\electronvolt}$.
      Results for the slices $\SI{45}{\giga\electronvolt}<M_{\ell\bar\ell}<\SI{60}{\giga\electronvolt}$ and $\SI{45}{\giga\electronvolt}<M_{\ell\bar\ell}<\SI{60}{\giga\electronvolt}$ can be found in figure~\ref{fig:cmsdy2d11_bins12}.}
    \label{fig:cmsdy2d11_bins3456}
\end{figure}
%-------------------------------------------------------------------------------

>>>>

In this section we demonstrate the capabilities of \textsc{PineAPPL} by
computing fast-in\-ter\-po\-la\-tion grids, accurate to NLO QCD and NLO QCD+EW,
for a representative set of processes in which EW corrections may have a
sizeable effect on the accuracy of the theoretical predictions.
In order to consider realistic kinematics for these
processes, we rely on the measurements that are commonly included in PDF
fits. Our aim is twofold. First, we want to validate the results
obtained with \textsc{PineAPPL}; second, we want to assess the
impact of the EW corrections for the specific experimental setups. We start by describing the
setup employed for the computation, then we discuss the features of each process, along with
the corresponding results.

@@

We generate each process by means of the Universal FeynRules Output
(UFO)~\cite{Degrande:2011ua} model {\tt loop\_qcd\_qed\_sm\_Gmu},
included as standard in {\sc MG5\_aMC}. It contains the UV and $R_2$
counterterms relevant to NLO QCD and EW corrections, the latter in the
$\overline{G}_\mu$ scheme. The model features five massless quark flavours,
sets the CKM matrix equal to the identity, and is compatible with the usage of
the complex mass (CM) scheme~\cite{Denner:1999gp,Denner:2005fg} for all massive particles, see
ref.~\cite{Frederix:2018nkq} for details. We use this scheme
for all processes that involve only massless particles in the final state.
The photon is always considered as part of the proton in the initial state and
of any hadronic jet produced in the final state: 
 in particular, PI contributions are consistently included at LO and NLO.\footnote{We employ the $\overline{G}_\mu$ scheme also for the QED coupling entering vertices involving initial-state photons, see section~4.3.3 of ref.\cite{Denner:2019vbn}.}
We use a PDF set that contains a photon PDF, namely
{\tt NNPDF31\_nlo\_as\_0118\_luxqed}~\cite{Bertone:2017bme}. We evaluate the PDF
uncertainty associated to the theoretical predictions a posteriori, that is,
we convolve the fast-interpolation grid generated with {\sc PineAPPL} with
each member in the PDF set, and we compute the associated standard deviation. Monte Carlo
weights are stored as Lagrange-interpolation grids.

@@

%-------------------------------------------------------------------------------
\begin{figure}[!p]
    \centering
    \includegraphics[width=0.46\textwidth]{figures/pineappl_CMSDY2D11_bin3}%
    \includegraphics[width=0.46\textwidth]{figures/pineappl_CMSDY2D11_bin4}\\
    \includegraphics[width=0.46\textwidth]{figures/pineappl_CMSDY2D11_bin5}%
    \includegraphics[width=0.46\textwidth]{figures/pineappl_CMSDY2D11_bin6}\\
    \caption{Same as figure~\ref{fig:atlaszhighmass49fb}, but for the CMS
      double-differential Drell--Yan lepton pair measurement at a
      centre-of-mass energy of \SI{7}{\tera\electronvolt}~\cite{Chatrchyan:2013tia}. Displayed are
      only two of the six invariant mass bins available, respectively below the
      Z-boson mass peak, $\SI{45}{\giga\electronvolt}<M_{\ell\bar\ell}<\SI{60}{\giga\electronvolt}$, on the Z-boson mass
      peak, $\SI{60}{\giga\electronvolt}<M_{\ell\bar\ell}<\SI{120}{\giga\electronvolt}$, above the mass peak,
      $\SI{120}{\giga\electronvolt}<M_{\ell\bar\ell}<\SI{200}{\giga\electronvolt}$, and at very high invariant masses,
      $\SI{200}{\giga\electronvolt}<M_{\ell\bar\ell}<\SI{1500}{\giga\electronvolt}$.
      Results for the slices $\SI{45}{\giga\electronvolt}<M_{\ell\bar\ell}<\SI{60}{\giga\electronvolt}$ and $\SI{45}{\giga\electronvolt}<M_{\ell\bar\ell}<\SI{60}{\giga\electronvolt}$ can be found in figure~\ref{fig:cmsdy2d11_bins12}.}
    \label{fig:cmsdy2d11_bins3456}
\end{figure}
%-------------------------------------------------------------------------------
